R Notebook

### 2020-09-15 simulation_studies/sim_study_small_model.Rmd
updated sim_studies

```{r}
# setting the stage
rm(list=ls())
options(scipen=999)
source("helper_plot_model.R")
source("helper_init_model.R")
source("helper_calc_mse.R")
source("helper_init_large_model.R")
library(asp20boost)
library(gamboostLSS)
library(microbenchmark)
```

@@

```{r}
# config 1 ----------------------
beta_1 <- c(0,1)
gamma_1 <- c(-3,2)

@@

```{r}
set.seed(1337)
mod_1 <- init_model(beta_1, gamma_1, 1000)$model
gradient_boost(mod_1)
mod_2 <- init_model(beta_2, gamma_2, 1000)$model
gradient_boost(mod_2)
mod_3 <- init_model(beta_3, gamma_3, 1000)$model
gradient_boost(mod_3)
mod_4 <- init_model(beta_4, gamma_4, 1000)$model
gradient_boost(mod_4)
```

```{r, cache=TRUE, fig.height = 8, fig.width = 11}
par(mfrow = c(2,2))
plot_model(beta_1, gamma_1, 1000, "Config 1")
plot_model(beta_2, gamma_2, 1000, "Config 2")
plot_model(beta_3, gamma_3, 1000, "Config 3")
plot_model(beta_4, gamma_4, 1000, "Config 4")
```
```{r}
result_df <- data.frame("Config" = c(1,1,2,2,3,3,4,4),
                        "beta_true" = c(beta_1, beta_2, beta_3, beta_4),
                        "beta est." = c(round(mod_1$beta, 3), round(mod_2$beta, 3), round(mod_3$beta, 3), round(mod_4$beta, 3)),
                        "gamma_true" = c(gamma_1, gamma_2, gamma_3, gamma_4),
                        "gamma est." = c(round(mod_1$gamma, 3), round(mod_2$gamma, 3), round(mod_3$gamma, 3), round(mod_4$gamma, 3))
                        )
result_df
```

@@

Configuration 2 is a somewhat naive parameter configuration, where all parameters are set to value 2. 
Estimating this model does not yield valid results.
The gradient_boost() function runs into numerical problems.
To understand why this happens we need to understand the model characteristics resulting from parameter configuration 2-2-2-2.
Here, the red arrow is of large size, while the blue arrow is hardly visible. 
This is due to the presence of huge standard deviations in response *y*, concealing changes in mean of *y.* 
In fact, the standard deviation at x = 1 is:

```{r}
exp(gamma_2[1] + gamma_2[2] * 1)
```
In contrast the true response mean values range from 0 to 2.
The change in response location (the signal), indicated by the blue arrow, perishes due to the high amount of noise caused by the high gamma-parameters. 
Resulting is a signal-noise ratio, that is impossible to estimate. 
Values for parameters gamma need to be chosen sensibly for simulating data. 
This is no straight-forward task, since the effect of *gamma* on the response standard deviation is not linear, but exponential. 
To check this in the two-dimensional case, the arrows in the model plot are of help.

@@

```{r}
mse_1 <- calc_MSE(beta_1, gamma_1, mod_1$beta, mod_1$gamma)
mse_3 <- calc_MSE(beta_3, gamma_3, mod_3$beta, mod_3$gamma)
cbind("config 1" = mse_1,
      "config_3" = mse_3)
```

@@

```{r}
mse_4 <- calc_MSE(beta_4, gamma_4, mod_4$beta, mod_4$gamma)
cbind("config 1" = mse_1,
      "config_3" = mse_3,
      "config 4" = mse_4)
```

>>>>

```{r}
# run all code chunks in this document
# do not display output of this codechunk
# DO display code
rm(list=ls())
options(scipen=999)
source("helper_plot_model.R")
source("helper_init_model.R")
source("helper_calc_mse.R")
source("helper_init_large_model.R")
library(asp20boost)
library(gamboostLSS)
library(microbenchmark)
```

@@

```{r}
# dont display code
# dont display output
# config 1 ----------------------
beta_1 <- c(0,1)
gamma_1 <- c(-3,2)

@@

```{r}
# display code
# dont display output
set.seed(1337)
mod_1 <- init_model(beta_1, gamma_1, 1000)$model
gradient_boost(mod_1)
mod_2 <- init_model(beta_2, gamma_2, 1000)$model
gradient_boost(mod_2)
mod_3 <- init_model(beta_3, gamma_3, 1000)$model
gradient_boost(mod_3)
mod_4 <- init_model(beta_4, gamma_4, 1000)$model
gradient_boost(mod_4)
```

```{r, cache=TRUE, fig.height = 8, fig.width = 11}
# display code
# display output
par(mfrow = c(2,2))
plot_model(beta_1, gamma_1, 1000, "Config 1")
plot_model(beta_2, gamma_2, 1000, "Config 2")
plot_model(beta_3, gamma_3, 1000, "Config 3")
plot_model(beta_4, gamma_4, 1000, "Config 4")
```
```{r}
# dont display code
# display output
result_df <- data.frame("Config" = c(1,1,2,2,3,3,4,4),
                        "beta_true" = c(beta_1, beta_2, beta_3, beta_4),
                        "beta est." = c(round(mod_1$beta, 3), round(mod_2$beta, 3), round(mod_3$beta, 3), round(mod_4$beta, 3)),
                        "gamma_true" = c(gamma_1, gamma_2, gamma_3, gamma_4),
                        "gamma est." = c(round(mod_1$gamma, 3), round(mod_2$gamma, 3), round(mod_3$gamma, 3), round(mod_4$gamma, 3))
                        )
result_df
```

@@

Configuration 2 is a somewhat naive parameter configuration, where all parameters are set to value 2. 
Estimating this model does not yield valid results.
The gradient_boost() function runs into numerical problems.
To understand why this happens we need to understand the model characteristics resulting from parameter configuration 2-2-2-2.
Here, the red arrow is of large size, while the blue arrow is hardly visible. 
This is due to the presence of huge standard deviations in response *y*, concealing changes in mean of *y.* 
In fact, the standard deviation at x = 1 is exp(2+2) = 54.60.
In contrast the true response mean values range from 0 to 2.
The change in response location (the signal), indicated by the blue arrow, perishes due to the high amount of noise caused by the high gamma-parameters. 
Resulting is a signal-noise ratio, that is impossible to estimate. 
Values for parameters gamma need to be chosen sensibly for simulating data. 
This is no straight-forward task, since the effect of *gamma* on the response standard deviation is not linear, but exponential. 
To check this in the two-dimensional case, the arrows in the model plot are of help.

@@

```{r}
# display code
# display output
mse_1 <- calc_MSE(beta_1, gamma_1, mod_1$beta, mod_1$gamma)
mse_3 <- calc_MSE(beta_3, gamma_3, mod_3$beta, mod_3$gamma)
cbind("config 1" = mse_1,
      "config_3" = mse_3)
```

@@

```{r}
# display code
# display output
mse_4 <- calc_MSE(beta_4, gamma_4, mod_4$beta, mod_4$gamma)
cbind("config 1" = mse_1,
      "config_3" = mse_3,
      "config 4" = mse_4)
```
